Real Estate

### 2022-12-05 src/anchor_city_real_estate.Rmd
calculate real estate score using anchor city occupancy rate

@@

```{r}
# aggregate yardi data into MSAs by cbsa
MSA_realEstate_data <- city_realEstate_data %>% 
  clean_names() %>% 
  rename(avg_rent2022 = x2022_rent) %>% 
   # calculate change in rent
  mutate(rent_change = (avg_rent2022 - average_rent_2021)/average_rent_2021 * 100) %>% 
  group_by(cbsa) %>% 
  # calculate average values for MSAs with more than one city
  summarize(avg_rent_2021 = mean(average_rent_2021, na.rm = TRUE),
            avg_rent_2022 = mean(avg_rent2022, na.rm = TRUE),
            avg_occupancy_21 = mean(percent_occupancy_2021, na.rm = TRUE),
            avg_rent_change = mean(rent_change, na.rm = TRUE))

@@

```{r}
# add MSA_median rent data
compliedMSA_realEstate<- select_MSAs %>% 
  left_join(MSA_realEstate_data) %>% 
    # calculate rent change
  mutate(med_rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100) %>% 
  # calculate income to homeprice
  mutate(house_price_2021 = as.numeric(house_price_2021)) %>% 
  mutate(income_to_homeprice = median_income/house_price_2021) %>% 
  # calculate rent to income ratio
  # multiply monthly rent x 12 for annual rent!
  mutate(med_rent_to_income = (median_rent2021*12)/median_income)

@@

```{r}
# normalize columns into new data table
n_MSA_realEstate <- compliedMSA_realEstate %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>%
  # normalize employment growth
  mutate(n_employ_change = (avg_emp_YoY - min(avg_emp_YoY)) / (max(avg_emp_YoY) - min(avg_emp_YoY))) %>% 
  # normalize rent to income
  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_med_rent_change = (med_rent_change - min(med_rent_change)) / (max(med_rent_change) - min(med_rent_change))) %>% 
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
  mutate(n_avg_occupancy = (avg_occupancy_21 - min(avg_occupancy_21, na.rm = TRUE)) / (max(avg_occupancy_21, na.rm = TRUE) - min(avg_occupancy_21, na.rm = TRUE))) %>% 
  # normalize rent change
  mutate(n_avg_rent_change = (avg_rent_change - min(avg_rent_change, na.rm = TRUE)) / (max(avg_rent_change, na.rm = TRUE) - min(avg_rent_change, na.rm = TRUE)))

@@

```{r}
# create downloadable data table
datatable(n_MSA_realEstate, extensions = 'Buttons', options = list(
    dom = 'Bfrtip',
    buttons = c('copy', 'csv', 'excel', 'pdf', 'print')
  )
)
```

```{r}
#write result to google sheet
# gs4_create("MSA_real_estate_score")
sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "n_MSA_realEstate")
```

>>>>

anchor_cities <- read_sheet("https://docs.google.com/spreadsheets/d/1leWRliw7cGGQTGIdBJzakjWZ4skQzSh04lFw_LfbtLU/edit#gid=0") %>%  clean_names()

@@

```{r}
# aggregate yardi data into MSAs by cbsa
city_realEstate_data <- city_realEstate_data %>% 
  clean_names() %>% 
  select(!x7) %>% 
  rename(avg_rent2022 = x2022_rent) %>% 
  rename(occ_rate2021 = percent_occupancy_2021)

anchor_cities_list <- unique(anchor_cities$city_msa)

anchor_city_data <- city_realEstate_data %>% 
  filter(city %in% c(anchor_cities_list))

@@

```{r}
# add MSA_median rent data
compliedMSA_realEstate<- select_MSAs %>% 
  left_join(anchor_city_data, by = "cbsa") %>% 
    # calculate rent change
  mutate(med_rent_change = (median_rent2022 - median_rent2021)/median_rent2021 * 100) %>% 
  # calculate income to homeprice
  mutate(house_price_2021 = as.numeric(house_price_2021)) %>% 
  mutate(income_to_homeprice = median_income/house_price_2021) %>% 
  # calculate rent to income ratio
  # multiply monthly rent x 12 for annual rent!
  mutate(med_rent_to_income = (median_rent2021*12)/median_income)

@@

```{r}
# normalize columns into new data table
n_MSA_realEstate <- compliedMSA_realEstate %>% 
  # normalize population data
  mutate(n_pop_change = (pop_percent_change20_21 - min(pop_percent_change20_21))/(max(pop_percent_change20_21)-min(pop_percent_change20_21))) %>%
  # normalize employment growth
  mutate(n_employ_change = (avg_emp_YoY - min(avg_emp_YoY)) / (max(avg_emp_YoY) - min(avg_emp_YoY))) %>% 
  # normalize rent to income
  mutate(n_med_rent_to_income = (med_rent_to_income- min(med_rent_to_income)) / (max(med_rent_to_income) - min(med_rent_to_income))) %>% 
  # normalize income to home price
  mutate(n_med_rent_change = (med_rent_change - min(med_rent_change)) / (max(med_rent_change) - min(med_rent_change))) %>% 
  mutate(n_income_to_homeprice = (income_to_homeprice - min(income_to_homeprice)) / (max(income_to_homeprice) - min(income_to_homeprice))) %>% 
  mutate(n_avg_occupancy = (occ_rate2021 - min(occ_rate2021)) / (max(occ_rate2021) - min(occ_rate2021)))

@@

```{r}
#write result to google sheet
#gs4_create("anchor_real_estate_score")
sheet_write(n_MSA_realEstate, ss = "https://docs.google.com/spreadsheets/d/1Ssmb7pUd4NFZocVp9C75SRMOV6tAJCCt0TjrCYBB8pY/edit?usp=sharing", sheet = "anchor_realEstate")
```
